# Choosing an effect size

The effect size is the back bone of meta-analysis as all other aspects of this book will branch out from initial effect size calculations that we make. 

When we talk about effect sizes we make the distinction between **measurement effect sizes** that are calculated based on means and standard deviations for a pair of treatment and control sites (there will be many of these in a meta-analysis) and a **summary effect size** (there will be one of these in the typical meta-analysis).  The summary effect size is the stastical aggregation of many measurement effect sizes.

There are many helpful texts and reviews that lay out the benefits and costs to using the many effect sizes that can be found in meta-analytic research.  One of the most popular packages for ecological meta-analysis, `metafor` allows for calculation of at least 15 effect sizes just by changing a few letters of you R code.  Please see these resources for more information on effect size options [@koricheva_handbook_2013, @schwarzer_meta-analysis_2015]. 

Within the `metafor` package the `escalc` function handles calculating the measurement-level effect size calculations. Essentially the function takes in pairs data from both treatment (abbreviated by a "t") and control sites (abbreviated with a "c"). The paired data typically represents sample means ($\bar{x_t}$ and $\bar{x_c}$), standard deviations ($s_t$ and $s_c$), and sample sizes ($n_t$ and $n_c$). Using these paired data, `escalc` will output a single effect size and measurement of variance for each row (i.e., paired measurement in your database). 

Here's how the paired data appear in our example database. Take a look at the table below and scroll to the right. You'll find colums that contain all of the quantitative information that is necessary for calculating an effect size for each pair of measurements or rows in the database. The column `mean_contol` has the mean species richness at uninvaded sites, the next column, `sample_size_control` provides the number of control study sites, however the authors defined "study site" within their article. You will be able to locate $\bar{x}$, $s$, and $n$ for every invasive species included in the database.  When we are missing some necessary metric, this is most often standard deviation, we can use methods to impute missing values, however, these will be covered in a later chapter.

```{r}
require(DT)
datatable(meta_analysis_data, fillContainer = TRUE)
```

Now, let's see how these calculations work with our example data.

## Standardized Mean Difference (aka Hedges' g)

The Standardized Mean Difference was one of the first effect sizes used in published meta-analyses [@hedges_statistical_1985] and it remains widely used in both ecological in medical meta- research [@crystal-ornelas_meta-analysis_nodate; @zavorka_negative_2018]. 

```{r}
SMD_effect_sizes <-
  escalc(
    "SMD",
    # Specify the effect size we want to calculate
    m1i = meta_analysis_data$mean_invaded,
    # provide means for group invaded sites
    n1i = meta_analysis_data$sample_size_invaded,
    # invaded site sample size
    sd1i = meta_analysis_data$SD_invaded,
    # invaded site SD
    m2i = meta_analysis_data$mean_control,
    # control mean
    n2i = meta_analysis_data$sample_size_control,
    # control sample size
    sd2i = meta_analysis_data$SD_control,
    # control SD
    data = meta_analysis_data
  )
```

## 
